---
title: "final_project_adv_1"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Hmisc)
library(tidyr)
library(corrplot)
library(BayesFactor)
library (BSDA)
library(MASS)
library(agricolae)
```


#Standard Biochemistry Profile
#https://wwwn.cdc.gov/nchs/nhanes/search/datapage.aspx?Component=Laboratory&CycleBeginYear=2017
#https://wwwn.cdc.gov/Nchs/Nhanes/2017-2018/DPQ_J.htm
#https://wwwn.cdc.gov/Nchs/Nhanes/2017-2018/BIOPRO_J.htm
#https://wwwn.cdc.gov/Nchs/Nhanes/2017-2018/MCQ_J.htm
#https://wwwn.cdc.gov/Nchs/Nhanes/2017-2018/HUQ_J.htm

The Data that is used for the project is taken from National Health and Nutrition Examination Survey's (NHANES) 2017-2018 Laboratory Data (Standard Biochemistry Profile) and Questionnaire Data (Mental Health - Depression Screener & Medical Conditions).
###2017-18

```{r}


bio_chem_full <- sasxport.get("BIOPRO_J_2017_18.XPT")
mental_health_data_1 <- mental_health_data <- sasxport.get("DPQ_J_2017_18.XPT")
hospital_utilization <- sasxport.get("HUQ_J_2017_18.XPT")

bio_chem_2017_18 <- merge(bio_chem_full, mental_health_data)


##combined data set, subsets of this data set will be use in this project to different type of statistical analysis 
bio_chem_2017_18 <- merge(bio_chem_2017_18, hospital_utilization)

bio_chem_2017_18<- bio_chem_2017_18[,c("lbxsatsi", "lbxsal","lbxscr", "lbxsgl" , "lbxsch" ,"lbxsir" ,"lbxstr", "lbxsapsi","lbxsassi","lbxsc3si","lbxsbu","lbxsclsi","lbxsck","lbxsgb","lbxsgtsi","lbxsldsi","lbxsossi","lbxsph","lbxsksi","lbxsnasi","lbxstb","lbxsca","lbxstp","lbxsua","dpq010","dpq020","dpq030","dpq040","dpq050","dpq060","dpq070","dpq090","huq090")]



```

###2015-16

```{r}


bio_chem_full <- sasxport.get("BIOPRO_J_2015_16.XPT")
mental_health_data_2 <- mental_health_data <- sasxport.get("DPQ_J_2015_16.XPT")
hospital_utilization <- sasxport.get("HUQ_J_2015_16.XPT")

bio_chem_2015_16 <- merge(bio_chem_full, mental_health_data)


##combined data set, subsets of this data set will be use in this project to different type of statistical analysis 
bio_chem_2015_16 <- merge(bio_chem_2015_16, hospital_utilization)

bio_chem_2015_16<- bio_chem_2015_16[,c("lbxsatsi",  "lbxsal","lbxscr", "lbxsgl" , "lbxsch" ,"lbxsir" ,"lbxstr", "lbxsapsi","lbxsassi","lbxsc3si","lbxsbu","lbxsclsi","lbxsck","lbxsgb","lbxsgtsi","lbxsldsi","lbxsossi","lbxsph","lbxsksi","lbxsnasi","lbxstb","lbxsca","lbxstp","lbxsua","dpq010","dpq020","dpq030","dpq040","dpq050","dpq060","dpq070","dpq090","huq090")]

```
###2013-14

```{r}


bio_chem_full <- sasxport.get("BIOPRO_J_2013_14.XPT")
mental_health_data_3 <- mental_health_data <- sasxport.get("DPQ_J_2013_14.XPT")
hospital_utilization <- sasxport.get("HUQ_J_2013_14.XPT")

bio_chem_2013_14 <- merge(bio_chem_full, mental_health_data)


##combined data set, subsets of this data set will be use in this project to different type of statistical analysis 
bio_chem_2013_14 <- merge(bio_chem_2013_14, hospital_utilization)

bio_chem_2013_14<- bio_chem_2013_14[,c("lbxsatsi", "lbxsal","lbxscr", "lbxsgl" , "lbxsch" ,"lbxsir" ,"lbxstr", "lbxsapsi","lbxsassi","lbxsc3si","lbxsbu","lbxsclsi","lbxsck","lbxsgb","lbxsgtsi","lbxsldsi","lbxsossi","lbxsph","lbxsksi","lbxsnasi","lbxstb","lbxsca","lbxstp","lbxsua","dpq010","dpq020","dpq030","dpq040","dpq050","dpq060","dpq070","dpq090","huq090")]

```
###2011-12

```{r}


bio_chem_full <- sasxport.get("BIOPRO_J_2011_12.XPT")
mental_health_data_4 <- mental_health_data <- sasxport.get("DPQ_J_2011_12.XPT")
hospital_utilization <- sasxport.get("HUQ_J_2011_12.XPT")

bio_chem_2011_12 <- merge(bio_chem_full, mental_health_data)


##combined data set, subsets of this data set will be use in this project to different type of statistical analysis 
bio_chem_2011_12 <- merge(bio_chem_2011_12, hospital_utilization)


bio_chem_2011_12<- bio_chem_2011_12[,c("lbxsatsi",  "lbxsal","lbxscr", "lbxsgl" , "lbxsch" ,"lbxsir" ,"lbxstr", "lbxsapsi","lbxsassi","lbxsc3si","lbxsbu","lbxsclsi","lbxsck","lbxsgb","lbxsgtsi","lbxsldsi","lbxsossi","lbxsph","lbxsksi","lbxsnasi","lbxstb","lbxsca","lbxstp","lbxsua","dpq010","dpq020","dpq030","dpq040","dpq050","dpq060","dpq070","dpq090","huq090")]
```




```{r}
bio_chem <- rbind(bio_chem_2017_18,bio_chem_2015_16)
bio_chem <- rbind(bio_chem,bio_chem_2013_14)
bio_chem <- rbind(bio_chem,bio_chem_2011_12)

bio_chem <- bio_chem %>% drop_na()


#yes hospital
#minor
nrow(subset(bio_chem, (huq090 == 1 & dpq020 == 1 )))
nrow(subset(bio_chem, (dpq020 == 1 )))
#major
nrow(subset(bio_chem, (huq090 == 1 & (dpq020 == 2)| (dpq020 == 3) )))

nrow(subset(bio_chem, (dpq020 == 2)| (dpq020 == 3) ))



#missing


# nrow(subset(bio_chem,  (dpq020 == 7)| (dpq020 == 9)| (dpq020 == ".") ))

```


The BIOPRO_J dataset has 41 variables initially. After removing all refrigerated serums and keeping only one type of measurement for a type of variable, I get 18 variables that represents body chemicals. I also added variables from mental health data (DPQ_J) that asks about 'Have little interest in doing things'(dpq010), 'Feeling down, depressed, or hopeless'(dpq020), and 'Trouble sleeping or sleeping too much' (dpq030). Also, rows that have NA values are dropped.

```{r}

bio_chem_trimmed<- bio_chem[,c("lbxsatsi","lbxsal","lbxscr", "lbxsgl" , "lbxsch" ,"lbxsir" ,"lbxstr", "lbxsapsi","lbxsassi","lbxsc3si","lbxsbu","lbxsclsi","lbxsck","lbxsgb","lbxsgtsi","lbxsldsi","lbxsossi","lbxsph","lbxsksi","lbxsnasi","lbxstb","lbxsca","lbxstp","lbxsua","dpq010","dpq020","dpq030","dpq040","dpq050","dpq060","dpq070","dpq090","huq090")]
bio_chem_trimmed <- bio_chem_trimmed %>% drop_na()

bio_chem_trimmed <- subset(bio_chem_trimmed, dpq010 != ".")
bio_chem_trimmed <- subset(bio_chem_trimmed, dpq010 != "7")
bio_chem_trimmed <- subset(bio_chem_trimmed, dpq010 != "9")

bio_chem_trimmed <- subset(bio_chem_trimmed, dpq020 != ".")
bio_chem_trimmed <- subset(bio_chem_trimmed, dpq020 != "7")
bio_chem_trimmed <- subset(bio_chem_trimmed, dpq020 != "9")

bio_chem_trimmed <- subset(bio_chem_trimmed, dpq030 != ".")
bio_chem_trimmed <- subset(bio_chem_trimmed, dpq030 != "7")
bio_chem_trimmed <- subset(bio_chem_trimmed, dpq030 != "9")

bio_chem_trimmed <- subset(bio_chem_trimmed, dpq040 != ".")
bio_chem_trimmed <- subset(bio_chem_trimmed, dpq040 != "7")
bio_chem_trimmed <- subset(bio_chem_trimmed, dpq040 != "9")

bio_chem_trimmed <- subset(bio_chem_trimmed, dpq050 != ".")
bio_chem_trimmed <- subset(bio_chem_trimmed, dpq050 != "7")
bio_chem_trimmed <- subset(bio_chem_trimmed, dpq050 != "9")

bio_chem_trimmed <- subset(bio_chem_trimmed, dpq060 != ".")
bio_chem_trimmed <- subset(bio_chem_trimmed, dpq060 != "7")
bio_chem_trimmed <- subset(bio_chem_trimmed, dpq060 != "9")

bio_chem_trimmed <- subset(bio_chem_trimmed, dpq070 != ".")
bio_chem_trimmed <- subset(bio_chem_trimmed, dpq070 != "7")
bio_chem_trimmed <- subset(bio_chem_trimmed, dpq070 != "9")

bio_chem_trimmed <- subset(bio_chem_trimmed, dpq090 != ".")
bio_chem_trimmed <- subset(bio_chem_trimmed, dpq090 != "7")
bio_chem_trimmed <- subset(bio_chem_trimmed, dpq090 != "9")

bio_chem_trimmed <- subset(bio_chem_trimmed, huq090 != ".")
bio_chem_trimmed <- subset(bio_chem_trimmed, huq090 != "7")
bio_chem_trimmed <- subset(bio_chem_trimmed, huq090 != "9")

bio_chem_trimmed$dpq010[bio_chem_trimmed$dpq010 != 0] <- 1
bio_chem_trimmed$dpq030[bio_chem_trimmed$dpq030 != 0] <- 1
bio_chem_trimmed$dpq040[bio_chem_trimmed$dpq040 != 0] <- 1
bio_chem_trimmed$dpq050[bio_chem_trimmed$dpq050 != 0] <- 1
bio_chem_trimmed$dpq060[bio_chem_trimmed$dpq060 != 0] <- 1
bio_chem_trimmed$dpq070[bio_chem_trimmed$dpq070 != 0] <- 1
bio_chem_trimmed$dpq090[bio_chem_trimmed$dpq090 != 0] <- 1



bio_chem_trimmed$huq090[bio_chem_trimmed$huq090 == 2] <- 0

bio_chem_trimmed <- bio_chem_trimmed %>% drop_na()
head(bio_chem_trimmed)
 
  #,"dpq010","dpq020","dpq030","dpq040","dpq050","dpq060","dpq070","dpq080","dpq090","dpq100"
```

```{r}
# model_dpq020 <- lm( dpq020 ~ 0+lbxsatsi+lbxsapsi+lbxsassi+lbxsc3si+lbxsbu+lbxsclsi+lbxsck+lbxsgb+lbxsgtsi+lbxsldsi+lbxsossi+lbxsph+lbxsksi+lbxsnasi+lbxstb+lbxsca+lbxstp+lbxsua+dpq010+dpq030+dpq040+dpq050+dpq060+dpq070+dpq090+huq090,data=bio_chem_trimmed)
# summary(model_dpq020)
# stepAIC(model_dpq020, direction='both', k = log(nrow(bio_chem_trimmed)))

```

```{r}
bio_chem_trimmed_all <- bio_chem_trimmed





bio_chem_trimmed_all$dpq020[bio_chem_trimmed_all$dpq020 != 0] <- 1

bio_chem_trimmed_all <- bio_chem_trimmed_all[c("lbxsatsi", "lbxsal","lbxscr", "lbxsgl" , "lbxsch" ,"lbxsir" ,"lbxstr", "lbxsapsi","lbxsassi","lbxsc3si","lbxsbu","lbxsclsi","lbxsck","lbxsgb","lbxsgtsi","lbxsldsi","lbxsossi","lbxsph","lbxsksi","lbxsnasi","lbxstb","lbxsca","lbxstp","lbxsua","dpq020")]


head(bio_chem_trimmed_all)


write.csv(bio_chem_trimmed_all,"bio_chem_trimmed_chem_mental.csv")

```

```{r}
library(synthpop)
nrow(subset(bio_chem_trimmed_all, dpq020 == "1"))
nrow(subset(bio_chem_trimmed_all, dpq020 == "0"))
df_pos <- subset(bio_chem_trimmed_all, dpq020 == "1")
h_dat <- syn(df_pos, k = 9900, method = "parametric", seed = 1)
write.syn(h_dat,file = "df_train_pos_syn.csv", filetype = "csv")


#synthtic data will be added manually in every case

```
```{r}
df_train_pos_syn <- read.csv("bio_chem_trimmed_chem_mental_synthetic.csv", header=TRUE) 
#df_train_pos_syn <- df_train_pos_syn[-1]
#bio_chem_trimmed_all <-df_train_pos_syn
```

##minor



```{r}
bio_chem_trimmed_normal <- bio_chem_trimmed
bio_chem_trimmed_normal <- subset(bio_chem_trimmed_normal, dpq020 != ".")
bio_chem_trimmed_normal <- subset(bio_chem_trimmed_normal, dpq020 != "7")
bio_chem_trimmed_normal <- subset(bio_chem_trimmed_normal, dpq020 != "9")
bio_chem_trimmed_normal <- subset(bio_chem_trimmed_normal, dpq020 != "2")
bio_chem_trimmed_normal <- subset(bio_chem_trimmed_normal, dpq020 != "3")

bio_chem_trimmed_normal$dpq020[bio_chem_trimmed_normal$dpq020 != 0] <- 1


bio_chem_trimmed_normal <- bio_chem_trimmed_normal %>% drop_na()
head(bio_chem_trimmed_normal)

bio_chem_trimmed_normal <- bio_chem_trimmed_normal[c("lbxsatsi", "lbxsal","lbxscr", "lbxsgl" , "lbxsch" ,"lbxsir" ,"lbxstr", "lbxsapsi","lbxsassi","lbxsc3si","lbxsbu","lbxsclsi","lbxsck","lbxsgb","lbxsgtsi","lbxsldsi","lbxsossi","lbxsph","lbxsksi","lbxsnasi","lbxstb","lbxsca","lbxstp","lbxsua","dpq020")]

write.csv(bio_chem_trimmed_normal,"bio_chem_trimmed_chem_mental_normal.csv")

```

```{r}


nrow(subset(bio_chem_trimmed_normal, dpq020 == "1"))
nrow(subset(bio_chem_trimmed_normal, dpq020 == "0"))
df_pos <- subset(bio_chem_trimmed_normal, dpq020 == "1")
h_dat <- syn(df_pos, k = 11000, method = "parametric", seed = 1)
write.syn(h_dat,file = "df_train_pos_syn_normal.csv", filetype = "csv")
```
##major





```{r}
bio_chem_trimmed_heavyl <- bio_chem_trimmed
bio_chem_trimmed_heavyl <- subset(bio_chem_trimmed_heavyl, dpq020 != ".")
bio_chem_trimmed_heavyl <- subset(bio_chem_trimmed_heavyl, dpq020 != "7")
bio_chem_trimmed_heavyl <- subset(bio_chem_trimmed_heavyl, dpq020 != "9")
bio_chem_trimmed_heavyl <- subset(bio_chem_trimmed_heavyl, dpq020 != "1")


bio_chem_trimmed_heavyl$dpq020[bio_chem_trimmed_heavyl$dpq020 != 0] <- 1


bio_chem_trimmed_heavyl <- bio_chem_trimmed_heavyl %>% drop_na()


bio_chem_trimmed_heavyl <- bio_chem_trimmed_heavyl[c("lbxsatsi", "lbxsal","lbxscr", "lbxsgl" , "lbxsch" ,"lbxsir" ,"lbxstr", "lbxsapsi","lbxsassi","lbxsc3si","lbxsbu","lbxsclsi","lbxsck","lbxsgb","lbxsgtsi","lbxsldsi","lbxsossi","lbxsph","lbxsksi","lbxsnasi","lbxstb","lbxsca","lbxstp","lbxsua","dpq020")]
head(bio_chem_trimmed_heavyl)




write.csv(bio_chem_trimmed_heavyl,"bio_chem_trimmed_chem_mental_heavy.csv")




```


```{r}

nrow(subset(bio_chem_trimmed_heavyl, dpq020 == "1"))
nrow(subset(bio_chem_trimmed_heavyl, dpq020 == "0"))
df_pos <- subset(bio_chem_trimmed_heavyl, dpq020 == "1")
h_dat <- syn(df_pos, k = 12500, method = "parametric", seed = 1)
write.syn(h_dat,file = "df_train_pos_syn_heavy.csv", filetype = "csv")

```